Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-03-01 paper/paper.tex
organizing plots + 4.3.3

\begin{proposition}
  The minimization algorithm reports rules that capture only the
  positive traffic.
\end{proposition}
\begin{proof}
  Every rule created during the search captures the positive malicious
  traffic as every rule contains a proper subset of the options from
  the original rule and all those options are satisfied by the
  malicious traffic. By definition of function $\mathit{isPlausible}$,
  these rules also capture \emph{only} the positive traffic. If
  negative traffic is captured, a rule is not added to $\mathcal R'$;
  consequently, it cannot be an element of $\mathcal O$ and reported
  on the output.
\end{proof}

@@

\begin{figure*}[t!]

@@

  \begin{subfigure}{.5\textwidth}

@@

    \scalebox{1.2}{

@@

            bar width=6pt,
            scaled ticks=false,
            tick label style={/pgf/number format/fixed},

@@

            enlargelimits=0.15,
            legend style={at={(0.5,-0.15)},
              anchor=north,legend columns=-1},

@@

@@

          xticklabel style={rotate=45},          

@@

  \vspace{-3ex}  
  \caption{\label{fig:distribution-strings}Strings used as parameters of content options.}

@@

%\vspace{1ex}
%\\
\begin{subfigure}{.5\textwidth}

@@

@@

          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={http\_method, http\_uri, http\_user\_agent, http\_header, http\_host, http\_connection, http\_request\_line},

@@

@@

        \addplot coordinates {(http\_method,3821) (http\_uri,10111) (http\_user\_agent,1022) (http\_header,4089) (http\_host,224) (http\_connection,29) (http\_request\_line,11)};

@@

  \caption{\label{fig:distribution-content_modifiers}Content modifiers
    frequency.}

@@

%\vspace{1ex}

@@

\begin{subfigure}{.5\textwidth}

@@

          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},

@@

          xticklabel style={rotate=45},          

@@

  \vspace{-3ex}  
  \caption{\label{fig:distribution-strings}Option Frequency.}

@@

\end{figure*}

\subsubsection{Number of content options.} We observed
that \percContentOptions\ of the rules from the ruleset we are
analyzing contain content options, but the distribution of those
contents per rule is not uniform. Figure~\ref{} shows the
corresponding distribution. Given the importance of this kind of
option to the definition of rules, we defined a heuristic function
focused on the typical number of content options in a rule. We
calculate the contribution of this function as described on previous
Section~\ref{sec:heuristic-size-of-rule}.

@@

\Mar{------------------------------------------- estou aqui}
\subsubsection{Popularity of content strings.} 
This heuristic weighs higher rules that use popular strings in content
options. For a given rule \tname{} takes the arithmetic mean of the
popularity scores of each string used in contents in that rule. More
precisely, we calculate the contribution of content rarity with the
following formula:

\[\frac{\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}}{N}\]

\noindent
, where the term $\mathit{content_i}$ denotes the $i$-th xcontent in a
given rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$-th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the ten most popular strings in our rule data set. To illustrate this
function, let us consider a rule with two content options, using the
strings ``GET'' and ``ID'', respectively. The contribution of that
rule for the computation of similarity score would be 0.601
(=(1+.203)/2).

\subsubsection{Popularity of content modifiers.} As previously
mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata
rules involve content options. Furthermore, of the 26K rules that use
content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
  desconsiderando HTTP? Esta parte eh generica.}. \Luc{Posso, mas nessa função nós consideramos apenas os modificadores http.} Content modifiers
serve to indicate what part of a network packet the string mentioned
in the content option should appear. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)
Figure~\ref{fig:http-header-example} illustrates possible content
options used to match against certain strings that appear in an HTTP
message. A check (resp., cross) mark indicates a match (resp.,
unmatch).

>>>>

\begin{proposition}
  \vspace{-1ex}  
  The minimization algorithm reports rules that capture only the
  positive traffic.
\end{proposition}
\begin{proof}
  \vspace{-1.5ex}
  Every rule created during the search captures the positive malicious
  traffic as every rule contains a proper subset of the options from
  the original rule and all those options are satisfied by the
  malicious traffic. By definition of function $\mathit{isPlausible}$,
  these rules also capture \emph{only} the positive traffic. If
  negative traffic is captured, a rule is not added to $\mathcal R'$;
  consequently, it cannot be an element of $\mathcal O$ and reported
  on the output.
  \vspace{-1ex}    
\end{proof}

@@

\begin{figure}[t!]

@@

  \begin{subfigure}{.25\textwidth}

@@

    \scalebox{0.8}{

@@

          bar width=4pt,
          scaled ticks=true,
          ymajorticks=false,

@@

          enlargelimits=0.08,
%          enlargelimits=false,
%          axis equal,
          clip=false,
          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          

@@

%          xticklabel style={rotate=45},          

@@

          xticklabel style={rotate=35},          

@@

  \vspace{-2ex}  
  \caption{\label{fig:distribution-strings}Strings used in content options.}

@@

\begin{subfigure}{.25\textwidth}

@@

          bar width=4pt,
          scaled ticks=true,
          ymajorticks=false,

@@

          enlargelimits=0.08,
%          enlargelimits=false,
%          axis equal,
          clip=false,
          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          

@@

          symbolic x coords={method, uri, user\_agent, header, host, connection, request\_line},          

@@

        \addplot coordinates {(method,3821) (uri,10111) (user\_agent,1022) (header,4089) (host,224) (connection,29) (request\_line,11)};

@@

  \vspace{-2ex}
  \caption{\label{fig:distribution-content_modifiers}Http content modifiers.}

@@

@@

\begin{subfigure}{.2\textwidth}

@@

          enlargelimits=0.08,
%          enlargelimits=false,
%          axis equal,
          clip=false,
          legend style={at={(0.5,-0.15)},anchor=north,legend columns=-1},          
          xticklabel style={rotate=35},          

@@

@@

  \vspace{-5ex}      
  \caption{\label{fig:distribution-strings}Options.}
%  \vspace{-1ex}    

@@

\end{figure}

\subsubsection{Number of content options.} We observed
that \percContentOptions\ of the rules from the ruleset we are
analyzing contain content options, but the distribution of those
contents per rule is not uniform. Figure~\ref{fig:distribution-contents} shows the
corresponding distribution. Given the importance of this kind of
option to the definition of rules, we defined a heuristic function
focused on the typical number of content options in a rule. We
calculate the contribution of this function as described on previous
Section~\ref{sec:heuristic-size-of-rule}.

@@

\subsubsection{Popularity of content strings.} 
The heuristic described in this section leverages content rarity to
discriminate rules---it weighs higher (respectively, lower) rules that
use popular (respectively, unpopular) strings in content options. The
score of a rule $r$ on this function is given by the arithmetic mean
of the popularity scores of each string used in contents in $r$. More
precisely, the contribution of content rarity is given by the
following formula:

\vspace{-2.5ex}
\[\frac{\sum_{i=1}^{N}\frac{\mathit{content\_frequency[content_i]}}{\mathit{max(content\_frequency)}}}{N}\]
\vspace{-2.5ex}

\noindent
, where $\mathit{content_i}$ denotes the $i$-th content in a given
rule, the term $\mathit{content\_frequency[\mathit{content_i}]}$
denotes the frequency of the string used in the $i$-th content, the
term $\mathit{max(content\_frequency)}$ denotes the maximum frequency
observed across all strings appearing as content parameters, and $N$
is the number of content options in a given
rule. Figure~\ref{fig:distribution-strings} shows the distribution of
the ten most popular strings in our ruleset. To illustrate this
function, let us consider a rule with two content options, using the
strings ``GET'' and ``ID'', respectively. The contribution of that
rule for the computation of similarity score would be 0.601
(=(1+.203)/2). The relatively high weight reflects the fact that
``GET'' and ``ID'' are popular contents.

\Mar{----------------------------------------------------------- estou
aqui}
\subsubsection{Popularity of content modifiers.} As previously
mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata
rules involve content options. Furthermore, of the 26K rules that use
content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
  desconsiderando HTTP? Esta parte eh generica.}. \Luc{Posso, mas nessa função nós consideramos apenas os modificadores http.} Content modifiers
serve to indicate what part of a network packet the string mentioned
in the content option should appear. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)
Figure~\ref{fig:http-header-example} illustrates possible content
options used to match against certain strings that appear in an HTTP
message. A check (resp., cross) mark indicates a match (resp.,
unmatch).
